Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-18 src/tex/ms.tex
Update

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Parameter space with mass ratio $q$ against post-newtonian spin parameter $\chi_{\mathrm{PN}}$. Orange: Training waveforms; Blue: Testing waveforms}
	\label{fig:intrin_space}
\end{figure}

@@

To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
from the SXS catalog in this study. We specifically select waveforms with 
negligible eccentricity (${e<2\times10^{-3}}$) and precession 
(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
parameters of the chosen test waveforms fall within the parameter space defined 
by the training waveforms. This suggests that these test waveforms can be used 
for direct comparison with the original model.

To illustrate the improvement on an individual waveform level, we compare a
particular waveform before and after optimization in Fig. \ref{fig:0154}.
Compared to the original IMRPhenomD waveform, we can see the optimized waveform
has smaller residual from NR waveform both in amplitude and phase, particularly
in the inspiral region, where the amplitude displays a $50\%$ reduction in
error. For a fair comparison, we selected one of the testing waveforms from the
catalog presented in \citep{khan2016frequency}.

To quantify the effect of optimization over the entire dataset, We evaluate the
mismatch of all testing waveforms using a constant PSD weighted loss function,
$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
\ref{fig:q148}. The peak of the distribution has shifted towards a lower
mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
improvement with a 22.9\% decrease in the median. Even though the distribution
lacks a clear peak due to the problem discussed in Section
\ref{subsec:optimization}, both distributions show observable improvement.

@@

%\kw{Which plot is this paragraph referring to?}
Applying the same methods, we will see that the distributions of mismatches
calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
unweighted mismatch.
The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
This result is expected, as the IMRPhenomD model was initially constructed and
fitted using the {\zdethp} weighted mismatch. Consequently, the model is
anticipated to closely fit the NR waveforms while incorporating the influence of
the {\zdethp} PSD, rather than a constant PSD.

>>>>

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and testing waveforms in $q-\chi_{PN}$
	space. The blue dots represent the waveforms used in training and the orange
	stars represent the waveforms used in testing.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
	\label{fig:intrin_space}
\end{figure}

@@

To evaluate how well the optimization procedure generalize to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and the data provided in the SXS catalog for 536 NR waveforms. We select
waveforms that share the same part of the parameter space with the training
waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
shows how the training and testing waveforms are distributed in the
$q-\chi_{PN}$ space.

To illustrate the effect of optimiziation on an individual waveform level, we
compare the mismatch of a particular waveform before and after optimization with
respect to the NR waveform taken directly from the SXS catalog in Fig.
\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
optimized waveform has smaller residual from NR waveform both in amplitude and
phase, particularly in the inspiral region, where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected one of the testing
waveforms from the catalog presented in \citep{khan2016frequency}.

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

@@

Note that the IMRPhenomD model was initially constructed and fitted using the
{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
methods using the {\zdethp} PSD exhibit superior improvement compared to the
unweighted mismatch. We find no significant qualitative and quantitative
difference between the two PSDs in the resulting distribution of mismatches.
